---
title: "Week 3 - Homework"
author: "STAT 420, Summer 2019, D. Unger"
date: ''
output:
  html_document: 
    toc: yes
  pdf_document: default
urlcolor: cyan
---

***

## Exercise 1 (Using `lm` for Inference)

For this exercise we will use the `cats` dataset from the `MASS` package. You should use `?cats` to learn about the background of this dataset.

**(a)** Fit the following simple linear regression model in `R`. Use heart weight as the response and body weight as the predictor. 

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Store the results in a variable called `cat_model`. Use a $t$ test to test the significance of the regression. Report the following:

- The null and alternative hypotheses
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.05$
- A conclusion in the context of the problem

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

```{r}
library(MASS)
cat_model = lm(Hwt ~ Bwt, data = cats)
summary(cat_model)$coefficients
```

```{r}
#test statistics
summary(cat_model)$coefficients[2, 3]
```
```{r}
#p value
summary(cat_model)$coefficients[2, 4]
```

 - $H_0: \beta_1 = 0$ 
 - $H_1: \beta_1 \neq 0$
 - Test statistics: t = 16.11939
 - P value = 6.969045e-34
 - We reject $H_0$ when $\alpha = 0.05$
 - The simple linear relationship exists between the heart weight variable and body weight variable.

**(b)** Calculate a 90% confidence interval for $\beta_1$. Give an interpretation of the interval in the context of the problem.

```{r}
confint(cat_model, "Bwt", level = 0.90)
```

The 90% confidence interval for $\beta_1$ is (3.619716, 4.448409). It means that we are 90% confident that for every 1 kg increase in body weight, the heart weight will increase in a range from 3.619716 g to 4.448409 g.
However, the confidence interval does not contain 0, therefore, $H_0$ might not hold true, and we need to reject it when $\alpha = 0.10$.

**(c)** Calculate a 99% confidence interval for $\beta_0$. Give an interpretation of the interval in the context of the problem.

```{r}
confint(cat_model, "(Intercept)", level = 0.99)
```

The 99% confidence interval for $\beta_0$ is (-2.164125, 1.4508). It means that we are 99% confident that when the body weight is 0 kg, the mean heart weight will be within the range from -2.164125 g to 1.4508 g. The confidence range contains 0, but it seems a bit lack of sense to consider the situation when body weight is 0 kg.

**(d)** Use a 99% confidence interval to estimate the mean heart weight for body weights of 2.1 and 2.8 kilograms. Which of the two intervals is wider? Why?

```{r}
new_bwt = data.frame(Bwt = c(2.1, 2.8))
a = predict(cat_model, newdata = new_bwt,
        interval = c("confidence"), level = 0.99)
a
```

When Bwt = 2.1 kg, 99% confidence interval is (7.599225, 8.630513).
When Bwt = 2.8 kg, 99% confidence interval is (10.618796, 11.258630).

```{r}
mean(cats$Bwt)
range(cats$Bwt)
```

The interval of the estimation of mean heart weight when body weight is 2.1 kg is larger because compared with 2.8 kg, 2.1 kg is farther away from the mean which is 2.723611 kg. Thus, the interval for the bwt = 2.1 kg has larger deviation, so it is larger. 

**(e)** Use a 99% prediction interval to predict the heart weight for body weights of 2.8 and 4.2 kilograms.

```{r}
new_bwt_2 = data.frame(Bwt = c(2.8, 4.2))
predict(cat_model, newdata = new_bwt_2,
        interval = c("prediction"), level = 0.99)
```

For bwt = 2.8 kg, the 99% prediction interval is (7.133247, 14.74418).
For bwt = 4.2 kg, the 99% prediction interval is (12.660883, 20.51192).

**(f)** Create a scatterplot of the data. Add the regression line, 90% confidence bands, and 90% prediction bands.

```{r}
bwt_grid = seq(min(cats$Bwt), max(cats$Bwt), by = 0.01)
hwt_ci_band = predict(cat_model, 
                       newdata = data.frame(Bwt = bwt_grid), 
                       interval = "confidence", level = 0.90)
hwt_pi_band = predict(cat_model, 
                       newdata = data.frame(Bwt = bwt_grid), 
                       interval = "prediction", level = 0.90) 
plot(Hwt ~ Bwt, data = cats,
     xlab = "Body weight (kg)",
     ylab = "Heart weight (g)",
     main = "Heart weight vs Body weight",
     pch  = 20,
     cex  = 2,
     col  = "grey",
     ylim = c(min(hwt_pi_band), max(hwt_pi_band)))
abline(cat_model, lwd = 5, col = "darkorange")

lines(bwt_grid, hwt_ci_band[,"lwr"], col = "dodgerblue", lwd = 3, lty = 2)
lines(bwt_grid, hwt_ci_band[,"upr"], col = "dodgerblue", lwd = 3, lty = 2)
lines(bwt_grid, hwt_pi_band[,"lwr"], col = "dodgerblue", lwd = 3, lty = 3)
lines(bwt_grid, hwt_pi_band[,"upr"], col = "dodgerblue", lwd = 3, lty = 3)
points(mean(cats$Bwt), mean(cats$Bwt), pch = "+", cex = 3)

```


**(g)** Use a $t$ test to test:

- $H_0: \beta_1 = 4$
- $H_1: \beta_1 \neq 4$

Report the following:

- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.05$

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

```{r}
summary(cat_model)$coefficients[2, ]
beta_1_hat = summary(cat_model)$coefficients[2, 1]
s_e = summary(cat_model)$coefficients[2, 2]
t = (beta_1_hat - 4) / s_e
n = length(resid(cat_model))
p_value = 2 * pt(abs(t), df = n - 2, lower.tail = FALSE )
t
p_value
```

 - test statistics: t = 0.1361084
 - p value: 0.8919283
 - We do not reject $H_0$ at $\alpha = 0.05$.

***

## Exercise 2 (More `lm` for Inference)

For this exercise we will use the `Ozone` dataset from the `mlbench` package. You should use `?Ozone` to learn about the background of this dataset. You may need to install the `mlbench` package. If you do so, do not include code to install the package in your `R` Markdown document.

For simplicity, we will re-perform the data cleaning done in the previous homework.

```{r}
data(Ozone, package = "mlbench")
Ozone = Ozone[, c(4, 6, 7, 8)]
colnames(Ozone) = c("ozone", "wind", "humidity", "temp")
Ozone = Ozone[complete.cases(Ozone), ]
```

**(a)** Fit the following simple linear regression model in `R`. Use the ozone measurement as the response and wind speed as the predictor. 

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Store the results in a variable called `ozone_wind_model`. Use a $t$ test to test the significance of the regression. Report the following:

- The null and alternative hypotheses
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.01$
- A conclusion in the context of the problem

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

```{r}
library(mlbench)
ozone_wind_model = lm(ozone ~ wind, data = Ozone)
summary(ozone_wind_model)$coefficients
```

```{r}
#t stats
summary(ozone_wind_model)$coefficients[2, 3]
```

```{r}
#p value
summary(ozone_wind_model)$coefficients[2, 4]
```

 - $H_0: \beta_1 = 0$ 
 - $H_1: \beta_1 \neq 0$
 - Test statistics: t = -0.2189811
 - P value = 0.8267954
 - We fail to reject $H_0$ when $\alpha = 0.01$
 - The simple linear relationship does not exist between ozone and wind speed.

**(b)** Fit the following simple linear regression model in `R`. Use the ozone measurement as the response and temperature as the predictor. 

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Store the results in a variable called `ozone_temp_model`. Use a $t$ test to test the significance of the regression. Report the following:

- The null and alternative hypotheses
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.01$
- A conclusion in the context of the problem

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

```{r}
ozone_temp_model = lm(ozone ~ temp, data = Ozone)
summary(ozone_temp_model)$coefficients
```
```{r}
#t stats
summary(ozone_temp_model)$coefficients[2, 3]
```
```{r}
#p value
summary(ozone_temp_model)$coefficients[2, 4]
```

 - $H_0: \beta_1 = 0$ 
 - $H_1: \beta_1 \neq 0$
 - Test statistics: t = 22.84896
 - P value = 8.153764e-71
 - We reject $H_0$ when $\alpha = 0.01$
 - The simple linear relationship exists between ozone and temperature.

***

## Exercise 3 (Simulating Sampling Distributions)

For this exercise we will simulate data from the following model:

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Where $\epsilon_i \sim N(0, \sigma^2).$ Also, the parameters are known to be:

- $\beta_0 = -5$
- $\beta_1 = 3.25$
- $\sigma^2 = 16$

We will use samples of size $n = 50$.

**(a)** Simulate this model $2000$ times. Each time use `lm()` to fit a simple linear regression model, then store the value of $\hat{\beta}_0$ and $\hat{\beta}_1$. Set a seed using **your** birthday before performing the simulation. Note, we are simulating the $x$ values once, and then they remain fixed for the remainder of the exercise.

```{r}
birthday = 19980121
set.seed(birthday)
n = 50
x = seq(0, 10, length = n)
```

```{r}
num_simu = 2000
beta_0 = -5
beta_1 = 3.25
sigma = 4
line = beta_0 + beta_1 * x

beta_hats = matrix(0, num_simu, 2)

for (i in 1:num_simu) {
  y = line + rnorm(n, mean = 0, sd = sigma)
  beta_hats[i, ] = coef(lm(y ~ x))
}
  
  beta_0_hats = beta_hats[ ,1]
  beta_1_hats = beta_hats[ ,2]
```


**(b)** Create a table that summarizes the results of the simulations. The table should have two columns, one for $\hat{\beta}_0$ and one for $\hat{\beta}_1$. The table should have four rows:

- A row for the true expected value given the known values of $x$
- A row for the mean of the simulated values
- A row for the true standard deviation given the known values of $x$
- A row for the standard deviation of the simulated values

```{r}
E_0 = -5
E_1 = 3.25
mean_0 = mean(beta_0_hats)
mean_0
mean_1 = mean(beta_1_hats)
mean_1

Sxx = sum((x - mean(x)) ^ 2)
true_sd_beta_0 = sigma * sqrt(1 / n + mean(x) ^ 2 / Sxx)
true_sd_beta_0
true_sd_beta_1 = sigma / sqrt(Sxx)
true_sd_beta_1

sd_0 = sd(beta_0_hats)
sd_0
sd_1 = sd(beta_1_hats)
sd_1

data_fra = data.frame(
  statistics = c("E[x]", "mean(x)", "SD(x)", "sd(x)"),
  beta_0_hats = c(E_0, mean_0, true_sd_beta_0, sd_0),
  beta_1_hats = c(E_1, mean_1, true_sd_beta_1, sd_1))

library(knitr)
kable(data_fra)
```


**(c)** Plot two histograms side-by-side:

- A histogram of your simulated values for $\hat{\beta}_0$. Add the normal curve for the true sampling distribution of $\hat{\beta}_0$.
- A histogram of your simulated values for $\hat{\beta}_1$. Add the normal curve for the true sampling distribution of $\hat{\beta}_1$.

```{r}
par(mfrow = c(1, 2))
hist(beta_0_hats, 
     xlab   = expression(hat(beta)[0]),
     prob = TRUE,
     breaks = 20,
     main = "",
     col    = "dodgerblue",
     border = "darkorange")
sd_beta0 = sigma * sqrt(1 / n + mean(x) ^ 2 / Sxx)
curve(dnorm(x, mean = beta_0, sd = sd_beta0) , add = TRUE, lwd = 3)


hist(beta_1_hats, prob = TRUE, breaks = 20, 
     xlab = expression(hat(beta)[1]), main = "", col    = "dodgerblue", border = "darkorange")
sd_beta1 = sigma / sqrt(Sxx)
curve(dnorm(x, mean = beta_1, sd = sd_beta1) , add = TRUE, lwd = 3)
```


***

## Exercise 4 (Simulating Confidence Intervals)

For this exercise we will simulate data from the following model:

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Where $\epsilon_i \sim N(0, \sigma^2).$ Also, the parameters are known to be:

- $\beta_0 = 5$
- $\beta_1 = 2$
- $\sigma^2 = 9$

We will use samples of size $n = 25$.

Our goal here is to use simulation to verify that the confidence intervals really do have their stated confidence level. Do **not** use the `confint()` function for this entire exercise.

**(a)** Simulate this model $2500$ times. Each time use `lm()` to fit a simple linear regression model, then store the value of $\hat{\beta}_1$ and $s_e$. Set a seed using **your** birthday before performing the simulation. Note, we are simulating the $x$ values once, and then they remain fixed for the remainder of the exercise.

```{r}
birthday = 19980121
set.seed(birthday)
n = 25
x = seq(0, 2.5, length = n)
```

```{r}
beta_0 = 5
beta_1 = 2
sigma = 3
line = beta_0 + beta_1 * x

num_simu = 2500
beta_1_hat = rep(0, 2500)
s_e = rep(0, 2500)

for (i in 1:num_simu) {
  y = line + rnorm(n, mean = 0, sd = sigma)
  beta_1_hat[i] = coef(lm(y ~ x))[2]
  s_e[i] = summary(lm(y ~ x))$sigma
}
```


**(b)** For each of the $\hat{\beta}_1$ that you simulated, calculate a 95% confidence interval. Store the lower limits in a vector `lower_95` and the upper limits in a vector `upper_95`. Some hints:

- You will need to use `qt()` to calculate the critical value, which will be the same for each interval.
- Remember that `x` is fixed, so $S_{xx}$ will be the same for each interval.
- You could, but do not need to write a `for` loop. Remember vectorized operations.

```{r}
crit = -qt(0.05 / 2, df = n - 2)
Sxx = sum((x - mean(x)) ^ 2)
lower_95 = beta_1_hat - crit * s_e / sqrt(Sxx)
upper_95 = beta_1_hat + crit * s_e / sqrt(Sxx)
```


**(c)** What proportion of these intervals contains the true value of $\beta_1$?

```{r}
mean(lower_95 < 2 & 2 < upper_95)
```

94.72% of these intervals contain the true value of $\beta_1$.

**(d)** Based on these intervals, what proportion of the simulations would reject the test $H_0: \beta_1 = 0$ vs $H_1: \beta_1 \neq 0$ at $\alpha = 0.05$?

```{r}
1 - mean(lower_95 < 0 & 0 < upper_95)
```

66.72% of the simulations would reject $H_0$.

**(e)** For each of the $\hat{\beta}_1$ that you simulated, calculate a 99% confidence interval. Store the lower limits in a vector `lower_99` and the upper limits in a vector `upper_99`.

```{r}
crit_99 = -qt(0.01 / 2, df = n - 2)
Sxx = sum((x - mean(x)) ^ 2)
lower_99 = beta_1_hat - crit_99 * s_e / sqrt(Sxx)
upper_99 = beta_1_hat + crit_99 * s_e / sqrt(Sxx)
```


**(f)** What proportion of these intervals contains the true value of $\beta_1$?

```{r}
mean(lower_99 < 2 & 2 < upper_99)
```

99.2% of the intervals contain the true value of $\beta_1$.

**(g)** Based on these intervals, what proportion of the simulations would reject the test $H_0: \beta_1 = 0$ vs $H_1: \beta_1 \neq 0$ at $\alpha = 0.01$?

```{r}
1 - mean(lower_99 < 0 & 0 < upper_99)
```

39.84% of the simulations would reject the test $H_o$.

***

## Exercise 5 (Prediction Intervals "without" `predict`)

Write a function named `calc_pred_int` that performs calculates prediction intervals:

$$
\hat{y}(x) \pm t_{\alpha/2, n - 2} \cdot s_e\sqrt{1 + \frac{1}{n}+\frac{(x-\bar{x})^2}{S_{xx}}}.
$$

for the linear model

$$
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i.
$$

**(a)** Write this function. You may use the `predict()` function, but you may **not** supply a value for the `level` argument of `predict()`. (You can certainly use `predict()` any way you would like in order to check your work.)

The function should take three inputs:

- `model`, a model object that is the result of fitting the SLR model with `lm()`
- `newdata`, a data frame with a single observation (row)
    - This data frame will need to have a variable (column) with the same name as the data used to fit `model`.
- `level`, the level (0.90, 0.95, etc) for the interval with a default value of `0.95`

The function should return a named vector with three elements:

- `estimate`, the midpoint of the interval
- `lower`, the lower bound of the interval
- `upper`, the upper bound of the interval

```{r}
calc_pred_int = function(model, newdata, level = 0.95){
  alpha = 1 - level
  n = length(resid(model))
  t = -qt(alpha / 2, df = n - 2)
    
  pre_int_95 = predict(model, newdata, interval = c("prediction"))
  t_95 = -qt(0.05 / 2, df = n-2)
  lower_95 = pre_int_95[2]
  upper_95 = pre_int_95[3]
  margin = (upper_95 - lower_95) / 2
  
  estimate = pre_int_95[1]
  se = margin / t_95
  
  lower_model = estimate - t * se
  upper_model = estimate + t * se
  results = c(estimate = estimate, lower = lower_model, upper = upper_model)
  
  results
}
```


**(b)** After writing the function, run this code:

```{r}
newcat_1 = data.frame(Bwt = 4.0)
calc_pred_int(cat_model, newcat_1)
```

**(c)** After writing the function, run this code:

```{r}
newcat_2 = data.frame(Bwt = 3.3)
calc_pred_int(cat_model, newcat_2, level = 0.99)
```


